---
title: "Prevalence of Community Gardens and Fridges in Areas of Food Insecurity in Buffalo, NY"
authors: Leah Bargnesi and Candace Kasahara
output:
  HTML
---

# Introduction

## Background
Food Insecurity is a serious issue that affects 10.5% of America as of 2019, according to Silva (2020). Additionally, due to systemic racism, Black and Hispanic Americans are twice as likely to face food insecurity compared to their white counterparts (Silva, 2020). Particularly, for Erie County, Feeding America (2020) states that 12% of residents are food insecure (110,390 individuals). Community efforts to support Buffalo and Erie County residents have taken the form of urban agriculture and community fridges. Besides providing food provisioning, studies have also shown that these efforts have many other social perks such as recreational benefits and a decrease in crime (Hess and Winner, 2007). The goal of this website is to show how these community efforts have improved food security over time. Demographic data will also be looked at. However, there are some limitations to this. There wasn't information on when each garden was founded, so some have no start dates. The other limitation is these maps right now can only show how food security may have improved and not the other benefits that were previously mentioned. Also, the COVID-19 pandemic did generally worsen food insecurity. Nevertheless, the idea is to show the importance of these community efforts.

## Problem / Question

How has access to community gardens and fridges improved food security in Buffalo, NY and surrounding areas?

# Materials and Methods

The goal of this website is to  make a map that highlights food security from 2009 to 2020 with points on the map where each garden and fridge is, and what year they were put in place. The data sets will be filtered out for Buffalo, NY and surrounding areas. The leaflet package will be utilized to make interactive maps. Data for food insecurity was requested and gathered from Feeding America, and locations of community gardens were requested and gathered from the University at Buffalo's Food Lab. Demographic data will also eventually be used and the package tidy census may be helpful for this.

# Results


## Loading Needed Packages
```{r, message=F, warning=F}
library(raster)
library(ggplot2)
library(ggmap)
library(rgdal)
library(rgl)
library(htmltools)
library(tidyverse)
library(leaflet)
library(kableExtra)
library(knitr)
library(dplyr)
library(vctrs)
library(tigris)
library(plotly)
library(maps)
library(tidycensus)
library(viridis)
library(scales)
library(data.table)
library(mapproj)
knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling
```


## Download Required Data
```{r}
com_url <- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/excel.csv"
senior_FS20_demog_url <- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/State_of_Senior_Housing_Datasets%20(1)/State%20of%20Senior%20Hunger%20in%202020%20Demographic.csv"
FS20_url<- read.csv("https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2022_2020-2019Data_ToShare.csv")
FS19_url<- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2021_2019Data_ToShare.csv"
FS18_url<- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2020_2018Data_ToShare.csv"
FS17_url<- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2019_2017Data_ToShare.csv"
FS16_url<- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2018_2016Data_ToShare.csv"
FS15_url<-"https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2017_2015Data_ToShare.csv"
FS14_url<-"https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2016_2014Data_ToShare.csv"
FS13_url<-"https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2015_2013Data_ToShare.csv"
FS12_url<-"https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2014_2012Data_ToShare.csv"
FS11_url<- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2013_2011Data_ToShare.csv"
FS10_url<-"https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2012_2010Data_ToShare.csv"
FS09_url<-"https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2011_2009Data_ToShare.csv"
```

## Download Food Insecurity Rate data for 2020
```{r}
FS2020_data<- FS20_url %>%
  rename(Food_Insecurity = X..of.Food.Insecure.Persons.Overall..1.Year., Food_Insecurity_Rate = Overall.Food.Insecurity.Rate..1.Year., County = County..State)



FS2020 <- subset(FS2020_data, select = c(State, County, Food_Insecurity_Rate, Food_Insecurity))
```

# U.S. Food Insecurity Choropleth Map
```{r}
USA <- counties(cb = TRUE)

USA_filter = USA %>% 
  mutate(FIPS = as.integer(GEOID)) %>%
  filter(!STATE_NAME %in% c("Alaska", "Hawaii", "Guam", "Commonwealth of the Northern Mariana Islands", "Puerto Rico", "American Samoa", "United States Virgin Islands"))
US_map = left_join(USA_filter,FS2020_data,by = "FIPS") %>% 
  mutate(Food_Insecurity_Rate2 = as.numeric(sub("%","",Food_Insecurity_Rate,fixed = T)))
my_breaks <- c(0, 10, 20, 30)


USA_filter = USA %>% 
  mutate(FIPS = as.integer(GEOID)) %>%
  filter(!STATE_NAME %in% c("Alaska", "Hawaii", "Guam", "Commonwealth of the Northern Mariana Islands", "Puerto Rico", "American Samoa", "United States Virgin Islands"))


US_map = left_join(USA_filter,FS2020_data,by = "FIPS") %>% 
  mutate(Food_Insecurity_Rate2 = as.numeric(sub("%","",Food_Insecurity_Rate,fixed = T)))

my_breaks <- c(0, 10, 20, 30)


ggplot() +
  geom_sf(data = US_map, color="transparent", size=0, aes(fill=Food_Insecurity_Rate2)) +
  labs(title = "Food Insecurity Rates in the U.S. (by County)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_gradientn(colours=rev(magma(6)),
                         name="Food_Insecurity_Rate",
                         na.value = "grey100",
                        breaks = my_breaks, labels = my_breaks)



```





## Clean all Required Data
```{r}
com_data<- read.csv(com_url)
senior_FS20_demog_data<- read.csv(senior_FS20_demog_url) #This is senior food insecurity demographic data for 2020 from Feeding America

FS20_data<- FS20_url%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year!=2019,Food.Insecurity.Rate = 
            as.numeric(sub("%","",Overall.Food.Insecurity.Rate..1.Year.)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
  
FS19_data<- read.csv(FS19_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2019,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2019.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)

FS18_data<- read.csv(FS18_url)%>%
  filter(X.1 =="Erie County, New York")%>%
   mutate(FIPS = as.character(Data.from.MMG.2020.are.NOT.directly.comparable.to.data.from.any.prior.MMG.study.due.to.methodological.changes.made.in.2020.), Year=2018,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X.2)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)

FS17_data<- read.csv(FS17_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2017,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2017.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)

FS16_data<- read.csv(FS16_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2016,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2016.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)

FS15_data<- read.csv(FS15_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2015,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2015.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)

FS14_data<- read.csv(FS14_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2014,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2014.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)

FS13_data<- read.csv(FS13_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2013,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2013.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)

FS12_data<- read.csv(FS12_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2012,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2012.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
 
 FS11_data<- read.csv(FS11_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2011,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2011.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
 

 FS10_data<- read.csv(FS10_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2010,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2010.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)


 FS09_data<- read.csv(FS09_url)%>%
  filter(County.Code=="ERIE", State.Name == "NY")%>%
   mutate(FIPS = as.character(FIPS),Year=2009,Food.Insecurity.Rate = 
            as.numeric(sub("%","",FI.Rate*100)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)

```



## Join Food Insecurity Datasets
```{r}
bind_all<- bind_rows(FS20_data, FS19_data, FS18_data, FS17_data, FS16_data, FS15_data, FS14_data, FS13_data, FS12_data, FS11_data, FS10_data, FS09_data)
bind_all[is.na(bind_all) | bind_all=="Inf"] = NA
```

# Results



## Urban farm and fridges locations

```{r}
com_data%>%
leaflet(options = 
        leafletOptions(minZoom = 5, dragging = TRUE))%>%
  addProviderTiles(provider = "Esri")%>%
  addMarkers(lng = com_data$long, lat = com_data$lat, popup = ~paste0(Name, "<br/>", Type, "<br/>", Address, "<br/>", Start.Year))%>%
  setView(lat=42.94991, lng=-78.89927, zoom=11)
  
  
```


## Plot of Food insecurity over time
```{r}
ggplot(bind_all, aes(Year, Food.Insecurity.Rate))+
  geom_point()+
  xlab("Year")+
  ylab("Rate of Food Insecurity")+
  ggtitle(label = "Rate of Food Insecurity from 2009-2020 for Erie County, New York")
  lm(formula = Food.Insecurity.Rate~Year, data= bind_all)
```



## Table of Food Insecurity and Food Provisioning Over Time
```{r}

kable(com_data)%>%# table of all gardens and fridges
kable_classic("striped", full_width = F)

#Table of gardens/fridges with start year after 2008 and food insecurity rate
gardens_w_year<- com_data%>%
  filter(Start.Year!="NA", Start.Year>2008)%>%
  mutate(Year = Start.Year)

table1<- merge(gardens_w_year, bind_all)

final_table <- table1[1:43, ]
final_table[9] <- cell_spec(final_table[[9]], color = "white", bold = T,
    background = spec_color(1:43, end = 0.9, option = "D", direction = 1, ))
kbl(final_table, escape = F, align = "c") %>%
  kable_classic("striped", full_width = F)


```
# Map of both food insecurity and gardens per year
```{r}
data(table1)

  # draw map
  map1<- map("county", region = "new york,erie")
  
  
 


```


# Conclusions

[~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.


# Sources

Hess, D., & Winner, L. 2007. Enhancing Justice and Sustainability at the Local Level: Affordable Policies for Urban Governments*. Local Environment, 12(4), 379–395. https://doi.org/10.1080/13549830701412489

Overall (all ages) Hunger & Poverty in the United States | Map the Meal Gap. 2022. Retrieved October 11, 2022, from https://map.feedingamerica.org

Silva, C. 2020, September 27. Food Insecurity In The U.S. By The Numbers. NPR. https://www.npr.org/2020/09/27/912486921/food-insecurity-in-the-u-s-by-the-numbers

Story Map Series. 2022 . https://suny-buffalo.maps.arcgis.com/apps/MapSeries/index.html?appid=7a3b2029ff374ac282a90573776ebffe.