---
title: "Prevalence of Community Gardens and Fridges in Areas of Food Insecurity in Buffalo, NY"
authors: Leah Bargnesi and Candice Kasahara
output:
  HTML
---

# Introduction

## Background
Food Insecurity is a serious issue that affects 10.5% of America as of 2019, according to Silva (2020). Additionally, due to systemic racism, Black and Hispanic Americans are twice as likely to face food insecurity compared to their white counterparts (Silva, 2020). Particularly, for Erie County, Feeding America (2020) states that 12% of residents are food insecure (110,390 individuals). Community efforts to support Buffalo and Erie County residents have taken the form of urban agriculture and community fridges. Besides providing food provisioning, studies have also shown that these efforts have many other social perks such as recreational benefits and a decrease in crime (Hess and Winner, 2007). The goal of this website is to show how these community efforts have improved food security over time. Demographic data will also be looked at. However, there are some limitations to this. There wasn't information on when each garden was founded, so some have no start dates. The other limitation is these maps right now can only show how food security may have improved and not the other benefits that were previously mentioned. Also, the COVID-19 pandemic did generally worsen food insecurity. Nevertheless, the idea is to show the importance of these community efforts.

## Problem / Question

How has access to community gardens and fridges improved food security in Buffalo, NY and surrounding areas?

# Materials and Methods

The goal of this website is to  make a map that highlights food security from 2009 to 2020 with points on the map where each garden and fridge is, and what year they were put in place. The data sets will be filtered out for Buffalo, NY and surrounding areas. The leaflet package will be utilized to make an interactive map of community garden locations. Data for food insecurity was requested and gathered from Feeding America, and locations of community gardens were requested and gathered from the University at Buffalo's Food Lab. First, food insecurity rates will be analyzed nationwide. Then food insecurity rates for Buffalo and the surrounding region will be examined for 2009-2020 to highlight the importance of community food provisioning.

# Results


## Loading Needed Packages
```{r, message=F, warning=F}
library(raster)
library(ggplot2)
library(ggmap)
library(rgdal)
library(rgl)
library(htmltools)
library(tidyverse)
library(leaflet)
library(kableExtra)
library(knitr)
library(dplyr)
library(vctrs)
library(tigris)
library(plotly)
library(maps)
library(tidycensus)
library(viridis)
library(scales)
library(data.table)
library(mapproj)
library(sf)
knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling
```


## Download Required Data For Food Insecurity and Community Gardens
```{r}
com_url <- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/excel.csv"
senior_FS20_demog_url <- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/State_of_Senior_Housing_Datasets%20(1)/State%20of%20Senior%20Hunger%20in%202020%20Demographic.csv"
FS20_url<- read.csv("https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2022_2020-2019Data_ToShare.csv")
FS19_url<- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2021_2019Data_ToShare.csv"
FS18_url<- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2020_2018Data_ToShare.csv"
FS17_url<- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2019_2017Data_ToShare.csv"
FS16_url<- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2018_2016Data_ToShare.csv"
FS15_url<-"https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2017_2015Data_ToShare.csv"
FS14_url<-"https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2016_2014Data_ToShare.csv"
FS13_url<-"https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2015_2013Data_ToShare.csv"
FS12_url<-"https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2014_2012Data_ToShare.csv"
FS11_url<- "https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2013_2011Data_ToShare.csv"
FS10_url<-"https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2012_2010Data_ToShare.csv"
FS09_url<-"https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2011_2009Data_ToShare.csv"
```

# U.S. Food Insecurity Map

To understand the degree of necessity for the implementation of urban agriculture and community fridges, we want to look at an overall map of the most recent food insecurity rates within the United States. With the information provided by this data, we want to compare it to poverty rates in the U.S. to determine whether they are significantly correlated. The data is from [Feeding America](https://www.feedingamerica.org/research/map-the-meal-gap/by-county).

## Food Insecurity Rate data for 2020
```{r}
FS20_url<- read.csv("https://raw.githubusercontent.com/geo511-2022/final_project-LeahBargnesi/master/data/Map_the_Meal_Gap_Data%20(1)/MMG2022_2020-2019Data_ToShare.csv")
FS2020_data<- FS20_url %>%
  rename(Food_Insecurity = X..of.Food.Insecure.Persons.Overall..1.Year., Food_Insecurity_Rate = Overall.Food.Insecurity.Rate..1.Year., County = County..State)

FS2020 <- subset(FS2020_data, select = c(State, County, Food_Insecurity_Rate, Food_Insecurity))
```

## U.S. Food Insecurity Choropleth Map 2020
```{r}
USA <- counties(cb = TRUE)

USA_filter = USA %>% 
  mutate(FIPS = as.integer(GEOID)) %>%
  filter(!STATE_NAME %in% c("Alaska", "Hawaii", "Guam", "Commonwealth of the Northern Mariana Islands", "Puerto Rico", "American Samoa", "United States Virgin Islands"))

US_map = left_join(USA_filter,FS2020_data,by = "FIPS") %>% 
  mutate(Food_Insecurity_Rate2 = as.numeric(sub("%","",Food_Insecurity_Rate,fixed = T)))
my_breaks <- c(0, 10, 20, 30)



US_map = left_join(USA_filter,FS2020_data,by = "FIPS") %>% 
  mutate(Food_Insecurity_Rate2 = as.numeric(sub("%","",Food_Insecurity_Rate,fixed = T)))

my_breaks <- c(0, 10, 20, 30)


ggplot() +
  geom_sf(data = US_map, color="transparent", size=0, aes(fill=Food_Insecurity_Rate2)) +
  labs(title = "Food Insecurity Rates in the U.S. (by County) in 2020", 
       caption = "Source: Map the Meal Gap 2022 - Feeding America") +
  theme_bw(13) +
  theme(plot.caption= element_text(size=7,
                                   color="blue",
                                   vjust = 5)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_gradientn(colours=rev(magma(6)),
                         name="Food Insecurity Rate",
                         na.value = "grey100",
                        breaks = my_breaks, labels = my_breaks)
  

```

Based on the choropleth map above, we see that most counties in the United States are experiencing food insecurity between 10-20%. There are some exceptions below 10%, especially in North Dakota. More noticeably, there are a significant number of counties that are facing food insecurity rates beyond 20%, most noticeably in South Dakota and spread throughout the southern portion of the U.S. As seen in the graph, food insecurity clearly threatens a significant portion of the U.S., which opens up the field of urban agriculture and community fridges to decrease some of the food insecurity being experienced.

# U.S. Poverty Line Map

After examining the food insecurity choropleth map above, we wanted to look at what some of the drivers of food insecurity to see which neighborhoods urban agriculture and community fridges should benefit or be implemented within. One driver we predict to correlate with food insecurity is the percent of those below the poverty line. According to the [ASPE](https://aspe.hhs.gov/topics/poverty-economic-mobility/poverty-guidelines/prior-hhs-poverty-guidelines-federal-register-references/2020-poverty-guidelines), the national 2020 poverty guideline was on average 12,760 dollars for one person in a household and 44,120 dollars for up to eight in one household. While the poverty guideline does vary per state, for this study, these averages will represent the entirety of the U.S. The data is from the United States Census, [SAIPE State and County Estimates for 2020](https://www.census.gov/data/datasets/2020/demo/saipe/2020-state-and-county.html).

## Poverty Percent Data for 2020
#```{r}
#US_poverty <- read.csv("Poverty.csv", skip = 5) [-1] %>%
 # mutate(FIPS = paste0(
  #  sprintf("%02d",State.FIPS.Code),
   # sprintf("%03d",County.FIPS.Code)),
   # Poverty_Percent=as.numeric(Poverty.Percent..All.Ages))
#```

## U.S. Poverty Percent Choropleth Map 2020
#```{r}

#USA_filtered = USA %>% 
  #mutate(FIPS = as.character(GEOID)) %>%
  filter(!STATE_NAME %in% c("Alaska", "Hawaii", "Guam", "Commonwealth of the Northern Mariana Islands", "Puerto Rico", "American Samoa", "United States Virgin Islands"))

Poverty_map = left_join(USA_filtered,US_poverty,by = "FIPS")
my_breaks <- c(0, 10, 20, 30)



Poverty_map = left_join(USA_filtered,US_poverty,by = "FIPS")

my_breaks <- c(0, 10, 20, 30)


ggplot() +
  geom_sf(data = Poverty_map, color="transparent", size=0, aes(fill=Poverty_Percent)) +
  labs(title = "Poverty Rates in the U.S. (by County) in 2020", 
       caption = "Source:  U.S. Census Bureau, Small Area Income
       and Poverty Estimates (SAIPE) Program") +
  theme_bw(13)+
  theme(plot.caption= element_text(size=7,
                                   color="blue",
                                   vjust = 5)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_gradientn(colours=rev(viridis(6)),
                         name="Poverty Rate",
                         na.value = "grey100",
                        breaks = my_breaks, labels = my_breaks)

#```

In this choropleth map, we see a similar trend as the food insecurity choropleth map. Overall, the poverty percent ranges from 5-20%. However, like the food insecurity map, some South Dakota counties appear to face a noticeably higher poverty rate, with over 30%, than other parts of the country. Similarly, the southern section of the U.S. shows between 20-30% poverty rate. Overall, we can see that there is high degree of correlation between poverty rate and percent of food insecurity.

# New York Poverty Percent vs. Food Insecurity

Seeing the visual correlation between the food insecurity and poverty percent choropleth maps, we deemed that it was significant enough to examine through a dot plot. We narrowed down our search to New York State since there are currently multiple initiatives that are combating food scarcity. If there is a noticeable correlation between food insecurity and percent poverty, then this information could be used to further push the implementation and funding for these efforts.

## Poverty Percent vs Food Insecurity Data
#```{r}
#Poverty <- US_poverty %>%
  #mutate(FIPS = as.integer(FIPS))
#Overall_map = left_join(Poverty,FS2020_data, by = "FIPS")

Overall2 <- Overall_map %>%
  filter(Postal.Code == "NY", Year == "2020") %>%
   mutate(Food_Insecurity_Rate2 = as.numeric(sub("%","",Food_Insecurity_Rate,fixed = T)))


Overall2 <- Overall_map %>%
  filter(Postal.Code == "NY", Year == "2020") %>%
   mutate(Food_Insecurity_Rate2 = as.numeric(sub("%","",Food_Insecurity_Rate,fixed = T)))


# Separating values needed for graphing
Overall3 <- subset(Overall2, select = c(FIPS, Food_Insecurity_Rate2, Poverty_Percent))
```

## Visualizing Poverty Percent and Food Insecurity Correlation
#```{r}
dot_plot <- ggplot(Overall3, aes(x = Poverty_Percent, y = Food_Insecurity_Rate2)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Poverty Percent vs. Food Insecurity in NYS Counties (2020)", x = "Poverty Percent", y = "Food Insecurity Rate") +
  theme_minimal(13) +
  theme(plot.title = element_text(hjust = 0.5))
dot_plot
```

In this dot plot, we see a clear correlation between poverty percent and food insecurity. As the poverty percent increases, food insecurity increases linearly as well. With this visualization, it is clear that efforts to reduce food insecurity could aid those who do not surpass the poverty line in New York State.
 
# Food Insecurity for Buffalo, New York and Food Provisioning Implementations

## Clean all Required Data for Food Insecurity for Erie County, New York

```{r}
com_data<- read.csv(com_url) #list of known community gardens/fridges, information on start year is known for some community gardens but not all.
#Data is filtered for Erie County New York and column names are changed for each year's data to match each other. Columns are also changed to character or numeric to match other year's data.
#2020 Food Insecurity for Erie County
FS20_data<- FS20_url%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year!=2019,Food.Insecurity.Rate = 
            as.numeric(sub("%","",Overall.Food.Insecurity.Rate..1.Year.)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
#2019 Food Insecurity for Erie County  
FS19_data<- read.csv(FS19_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2019,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2019.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
#2018 Food Insecurity for Erie County
FS18_data<- read.csv(FS18_url)%>%
  filter(X.1 =="Erie County, New York")%>%
   mutate(FIPS = as.character(Data.from.MMG.2020.are.NOT.directly.comparable.to.data.from.any.prior.MMG.study.due.to.methodological.changes.made.in.2020.), Year=2018,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X.2)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
#2017 Food Insecurity for Erie County
FS17_data<- read.csv(FS17_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2017,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2017.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
#2016 Food Insecurity for Erie County
FS16_data<- read.csv(FS16_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2016,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2016.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
#2015 Food Insecurity for Erie County
FS15_data<- read.csv(FS15_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2015,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2015.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
#2014 Food Insecurity for Erie County
FS14_data<- read.csv(FS14_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2014,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2014.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
#2013 Food Insecurity for Erie County
FS13_data<- read.csv(FS13_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2013,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2013.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
#2012 Food Insecurity for Erie County
FS12_data<- read.csv(FS12_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2012,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2012.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
 
#2011 Food Insecurity for Erie County
 FS11_data<- read.csv(FS11_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2011,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2011.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
 
 #2010 Food Insecurity for Erie County
 FS10_data<- read.csv(FS10_url)%>%
  filter(County..State=="Erie County, New York")%>%
   mutate(FIPS = as.character(FIPS),Year=2010,Food.Insecurity.Rate = 
            as.numeric(sub("%","",X2010.Food.Insecurity.Rate)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
 
 #2009 Food Insecurity for Erie County
 FS09_data<- read.csv(FS09_url)%>%
  filter(County.Code=="ERIE", State.Name == "NY")%>%
   mutate(FIPS = as.character(FIPS),Year=2009,Food.Insecurity.Rate = 
            as.numeric(sub("%","",FI.Rate*100)))%>%
   select(FIPS,Year,Food.Insecurity.Rate)
```


## Join Food Insecurity Datasets
Food Insecurity data for all years (2009-2020) are joined together in one data frame.
```{r}
bind_all<- bind_rows(FS20_data, FS19_data, FS18_data, FS17_data, FS16_data, FS15_data, FS14_data, FS13_data, FS12_data, FS11_data, FS10_data, FS09_data)
bind_all[is.na(bind_all) | bind_all=="Inf"] = NA
```


## Interactive map of Urban Farm and Fridges Locations
Markers for Erie County New York represent all known community gardens, urban farms,or fridges. Markers can be clicked on to view name, address, type of food provisioning, and start year if known.
```{r}
com_data%>%
leaflet(options = 
        leafletOptions(minZoom = 5, dragging = TRUE))%>%
  addProviderTiles(provider = "Esri")%>%
  addMarkers(lng = com_data$long, lat = com_data$lat, popup = ~paste0(Name, "<br/>", Type, "<br/>", Address, "<br/>", Start.Year))%>%
  setView(lat=42.94991, lng=-78.89927, zoom=11)
  
  
```


## Plot of Food insecurity over time
This graph depicts how food insecurity has changed from 2009-2020.
```{r}
ggplot(bind_all, aes(Year, Food.Insecurity.Rate))+
  geom_point()+
  xlab("Year")+
  ylab("Rate of Food Insecurity")+
  ggtitle(label = "Rate of Food Insecurity from 2009-2020 for Erie County, New York")
  lm(formula = Food.Insecurity.Rate~Year, data= bind_all)
```
You can see that food insecurity is on a general decline, though it did spike again in 2020 (height of the COVID-19 pandemic). Yet, food insecurity for 2020 still did not reach the highest recorded food insecurity rate for the years 2009-2020.


## Table of Food Insecurity and Food Provisioning Over Time
Two tables were made, so that you can easily view the amount of community gardens there are in Buffalo, New York. The first table lists all known food provisioning resources, and the second table shows only the food provisioning resources after 2008. This is done so that you can see how the community gardens/fridges that were implemented after 2008 may have impacted food insecurity rates. 
```{r}
kable(com_data)%>% # table of all gardens and fridges
kable_classic("striped", full_width = F)
#Table of gardens/fridges with start year after 2008 and food insecurity rate
gardens_w_year<- com_data%>%
  filter(Start.Year!="NA", Start.Year>2008)%>%
  mutate(Year = Start.Year)
table1<- merge(gardens_w_year, bind_all)
final_table <- table1[1:43, ]
final_table[9] <- cell_spec(final_table[[9]], color = "white", bold = T,
    background = spec_color(1:43, end = 0.9, option = "D", direction = 1, ))
kbl(final_table, escape = F, align = "c") %>%
  kable_classic("striped", full_width = F)
```

# Map of both food insecurity and gardens per year
The goal for this section is to show visually how food insecurity has improved as more gardens are implemented. The maps will be faceted by year and points will be added on map for each cumulative amount of gardens in that year. The map will be filled in by color that corresponds to food insecurity rates.
```{r}
data(table1)
  # draw map
  map1<- map("county", region = "new york,erie")
  
  
 
```


# Conclusions

Food insecurity is an issue in many regions. There are many factors that influence who will be most affected by food insecurity. One of those factors is individuals that live below the poverty line. Because there are little solutions implemented by the government, many times community efforts make the most impact. The goal of this website was to highlight how the grassroots implementation of community gardens, urban farms, and fridges in Erie County, New York have made an improvement in food insecurity. As, more food provisioning has increased in Erie County there has been a decrease in food insecurity rates. There was a bit of a spike in 2020 which is most likely due to the pandemic in which many people became unemployed or were unable to have access to healthy food due to illness or fear of contracting COVID-19. There also may be other factors that have led to the decrease of food insecurity in Erie County, New York. Nevertheless, we think this can give a general idea about how grassroots actions like community gardens are important and can make a difference for the resilience and sustainability of a community.


# Sources

Gundersen, C., Strayer, M., Dewey, A., Hake, M., & Engelhard, E. (2022). Map the Meal Gap 2022: An Analysis of County and Congressional District Food Insecurity and County Food Cost in the United States in 2020. Feeding America. 

Hess, D., & Winner, L. 2007. Enhancing Justice and Sustainability at the Local Level: Affordable Policies for Urban Governments*. Local Environment, 12(4), 379–395. https://doi.org/10.1080/13549830701412489

Overall (all ages) Hunger & Poverty in the United States | Map the Meal Gap. 2022. Retrieved October 11, 2022, from https://map.feedingamerica.org

Silva, C. 2020, September 27. Food Insecurity In The U.S. By The Numbers. NPR. https://www.npr.org/2020/09/27/912486921/food-insecurity-in-the-u-s-by-the-numbers

Story Map Series. 2022 . https://suny-buffalo.maps.arcgis.com/apps/MapSeries/index.html?appid=7a3b2029ff374ac282a90573776ebffe.

2020 Poverty and Median Household Income Estimates - Counties, States, and National

Source:  U.S. Census Bureau, Small Area Income and Poverty Estimates (SAIPE) Program

